---
title: "RFC_FirstExploration"
author: "Madlen Wilmes"
date: "6/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
library(usmap)
```



## Read human readable averaged scans
The data can be downloaded from the [gitlab of the Real Food Campaign](https://gitlab.com/our-sci/rfc-docs/-/tree/master/2019)

```{r ingest human readable avg scans}
dat <- read.csv("./rfc-docs/2019/completeDatasetHumanReadableMeans.csv")
str(dat)
# There are 127 measures in the (reduced!) data set!
# list all variables
colnames(dat)
head(dat[,1:10])
```

```{r basics}
dim(dat)
```

```{r summary}
summary(dat)
```

## Findings from `summary`
- variable "Unique.Planting" is entirely empty
- variable "Share.Code" contains only zeros
- free-text fields not very useful (e.g., does Quality.in.field:spotted mean  the same as Quality.in.field:pest?)
- should "quality.in.lab:rotten" even be considered?
- does "Brix" only measure sucrose?
- approximately >90% of samples do not have mineral data from crop or soil
- ~ 20% does not have juice scan data
- ~ 80% does not have seetness/taste information


## Questions from `summary`
- Do farm practices matter (e.g., hydroponic vs others)
- Does `Quality.in.field` matter? (e.g., 'pest')
- What does Quality.in.lab:yes mean?



```{r counts of per crop}
dat %>% group_by(Type) %>% summarize(counts = n()) %>% arrange(desc(counts))
```


```{r counts of varieties within crops}
dat %>% group_by(Type) %>% summarize(probes_count = n(), variety_counts = n_distinct(Variety), .groups='keep') %>% arrange(desc(variety_counts))
```

```{r completeness of source per variety}
dat %>% group_by(Type, Variety) %>% summarize(source_count = n_distinct(Source), .groups='keep')
```
Only few varieties have all three levels of source ("farm","farm_market", "store")

```{r distribution of source across produce Type}
dat %>% group_by(Source) %>% summarize(sample_count = n())
dat %>% group_by(Type, Source) %>% summarize(sample_count = n()) %>% 
  ggplot(aes(Type, sample_count)) +
  geom_bar(aes(fill = Source), position = "dodge", stat="identity")
```


```{r plot difference in antiox by source}
dat %>% 
  group_by(Type, Source) %>% 
  ggplot(aes(Type, Antioxidants)) +
  geom_boxplot(aes(fill = Source))
```


```{r stats difference by source}
fit1 <- aov(Antioxidants ~ Type:Source, data = dat)
summary(fit1)
TukeyHSD(fit1)
hist(residuals(fit1),
     col="darkgray")
```


```{r plot difference in Dry.Mass}
dat %>% 
  group_by(Type, Source) %>% 
  ggplot(aes(Type, Dry.Mass....x.100.)) +
  geom_boxplot(aes(fill = Source))
```


```{r plot difference in Crop..Potassium}
dat %>% 
  group_by(Type, Source) %>% 
  ggplot(aes(Type, Dry.Mass....x.100.)) +
  geom_boxplot(aes(fill = Source))
```



```{r sample contributions at state level}
dat %>% group_by(State, Type, Source) %>% summarize(sample_count = n())
```
